School closures and consolidations: the active travel and emission implications of reduced accessibility

### 2024-11-21 data-prep/02-travel-times-ODs.qmd
Added 2021 schools and census to the database. Started Manuscript changes.

```{r}
library(hamONdests)
data(Schools_2011_2016)
data(School_Catchments_2011_2016)
load(file = "data-inputs/Parcels/RED_LU_centroids_2011.Rdata")
load(file = "data-inputs/Parcels/RED_LU_centroids_2016.Rdata")
```

Format the destination points (ID, Long, Lat):
```{r}
Dest_all_schools_elem_2011 <- cbind(as.data.frame(Schools_2011_2016), st_coordinates(st_transform(Schools_2011_2016, crs = 4326))) %>% filter(Level == "Elementary" & (Year == "2011" | Year == "2011 and 2016")) %>%
  rename(lon = "X", lat = "Y", id = "SchoolID") %>%
  select(id, lon, lat)

Dest_all_schools_elem_2016 <- cbind(as.data.frame(Schools_2011_2016), st_coordinates(st_transform(Schools_2011_2016, crs = 4326))) %>% filter(Level == "Elementary" & (Year == "2016" | Year == "2011 and 2016")) %>%
  rename(lon = "X", lat = "Y", id = "SchoolID") %>%
  select(id, lon, lat)
```

Format the origin points (ID, Long, Lat):
```{r}
Orig_all_2011 <- cbind(RED_LU_centroids_2011,
                       st_coordinates(st_transform(RED_LU_centroids_2011 |> st_as_sf(), crs = 4326))) %>%
  rename(lon = X, lat = Y, id = ID) %>%
  select(id, lon, lat)

Orig_all_2016 <- cbind(RED_LU_centroids_2016,
                       st_coordinates(st_transform(RED_LU_centroids_2016 |> st_as_sf(), crs = 4326))) %>%
  rename(lon = X, lat = Y, id = ID) %>%
  select(id, lon, lat)
```

@@

@@

# Set Up R5 Routing
```{r build graph, include = FALSE}
r5_HAM <- setup_r5(data_path = r5_path, verbose = TRUE, overwrite = TRUE, temp_dir = TRUE)
```

@@

2011 Car travel times:
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2011_1)/chunksize)

@@

2016 car travel times:
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2016_1)/chunksize)

@@

2011 walk travel times:
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2011_1)/chunksize)

@@

2016 walk travel times:
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2016_1)/chunksize)

@@

>>>>

```{r}
# library(hamONdests)
# data(SCHOOLS)
# data(School_Catchments_2011_2016)
load(file="data-products/SCHOOLS.Rdata") #for this paper, the schools under analysis
load(file = "data-inputs/Parcels/RED_LU_centroids_2011.Rdata")
load(file = "data-inputs/Parcels/RED_LU_centroids_2016.Rdata")
load(file = "data-inputs/Parcels/RED_LU_centroids_2021.Rdata")
```

Format the destination points (ID, Long, Lat):
```{r}
Dest_all_schools_elem_2011 <- cbind(as.data.frame(SCHOOLS), st_coordinates(st_transform(SCHOOLS, crs = 4326))) %>% filter(Level == "Elementary" & (Year == "2011" | Year == "2011 and 2016")) %>%
  rename(lon = "X", lat = "Y", id = "SchoolID") %>%
  select(id, lon, lat)

Dest_all_schools_elem_2016 <- cbind(as.data.frame(SCHOOLS), st_coordinates(st_transform(SCHOOLS, crs = 4326))) %>% filter(Level == "Elementary" & (Year == "2016" | Year == "2011 and 2016")) %>%
  rename(lon = "X", lat = "Y", id = "SchoolID") %>%
  select(id, lon, lat)

Dest_all_schools_elem_2021 <- cbind(as.data.frame(SCHOOLS), st_coordinates(st_transform(SCHOOLS, crs = 4326))) %>% filter(Level == "Elementary" & (Year_2 == "2021" | Year_2 == "2016 and 2021")) %>%
  rename(lon = "X", lat = "Y", id = "SchoolID") %>%
  select(id, lon, lat)
```

Format the origin points (ID, Long, Lat):
```{r}
Orig_all_2011 <- cbind(RED_LU_centroids_2011,
                       st_coordinates(st_transform(RED_LU_centroids_2011 |> st_as_sf(), crs = 4326))) %>%
  rename(lon = X, lat = Y, id = ID) %>%
  dplyr::select(c(id, lon, lat))

Orig_all_2016 <- cbind(RED_LU_centroids_2016,
                       st_coordinates(st_transform(RED_LU_centroids_2016 |> st_as_sf(), crs = 4326))) %>%
  rename(lon = X, lat = Y, id = ID) %>%
  select(id, lon, lat)

Orig_all_2021 <- cbind(RED_LU_centroids_2021,
                       st_coordinates(st_transform(RED_LU_centroids_2021 |> st_as_sf(), crs = 4326))) %>%
  rename(lon = X, lat = Y, id = ID) %>%
  select(id, lon, lat)
```

@@

```{r}
Orig_all_2021_1 <- Orig_all_2021[1:36903,]
Orig_all_2021_2 <- Orig_all_2021[36904:73807,]
Orig_all_2021_3 <- Orig_all_2021[73808:110710,]
Orig_all_2021_4 <- Orig_all_2021[110711:143890,]
```

@@

# Set Up R5 Routing
```{r build graph, include = FALSE}
r5_HAM <- setup_r5(data_path = r5_path, verbose = TRUE,  temp_dir = TRUE)#, overwrite = TRUE)
```

@@

## 2011 Car travel times:
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2011_1)/chunksize)

@@

## 2016 car travel times:
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2016_1)/chunksize)

@@

## 2021 car travel times:
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2021_1)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(Orig_all_2021_1,
                          outdir = "data-inputs/Travel-times/Processing/Orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = Dest_all_schools_elem_2021,
                          mode = c("CAR"), 
                          departure_datetime = as.POSIXct(strptime("2021-03-01 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          max_trip_duration = 90)

save("output_OD_car1", file = "data-products/OD_tt_car_2021_1.Rdata")
rm("output_OD_car1")
```

```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2021_2)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(Orig_all_2021_2,
                          outdir = "data-inputs/Travel-times/Processing/Orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = Dest_all_schools_elem_2021,
                          mode = c("CAR"), 
                          departure_datetime = as.POSIXct(strptime("2021-03-01 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          max_trip_duration = 90)

save("output_OD_car2", file = "data-products/OD_tt_car_2021_2.Rdata")
rm("output_OD_car2")
```

```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2021_3)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(Orig_all_2021_3,
                          outdir = "data-inputs/Travel-times/Processing/Orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = Dest_all_schools_elem_2021,
                          mode = c("CAR"), 
                          departure_datetime = as.POSIXct(strptime("2021-03-01 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          max_trip_duration = 90)

save("output_OD_car3", file = "data-products/OD_tt_car_2021_3.Rdata")
rm("output_OD_car3")
```

```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2021_4)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(Orig_all_2021_4,
                          outdir = "data-inputs/Travel-times/Processing/Orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = Dest_all_schools_elem_2021,
                          mode = c("CAR"), 
                          departure_datetime = as.POSIXct(strptime("2021-03-01 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          max_trip_duration = 90)

save("output_OD_car4", file = "data-products/OD_tt_car_2021_4.Rdata")
rm("output_OD_car4")
```

## 2011 walk travel times:
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2011_1)/chunksize)

@@

## 2016 walk travel times:
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2016_1)/chunksize)

@@

## 2021 walk travel times:
```{r}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2021_1)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(Orig_all_2021_1,
                          outdir = "data-inputs/Travel-times/Processing/Orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = Dest_all_schools_elem_2021,
                          mode = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("2021-03-01 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          walk_speed = 3.6,
                          max_trip_duration = 30)

save("output_OD_walk_elem1", file = "data-products/OD_tt_walk_2021_1.Rdata")
rm("output_OD_walk_elem1")

# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2021_2)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(Orig_all_2021_2,
                          outdir = "data-inputs/Travel-times/Processing/Orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = Dest_all_schools_elem_2021,
                          mode = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("2021-03-01 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          walk_speed = 3.6,
                          max_trip_duration = 30)

save("output_OD_walk_elem2", file = "data-products/OD_tt_walk_2021_2.Rdata")
rm("output_OD_walk_elem2")

# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2021_3)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(Orig_all_2021_3,
                          outdir = "data-inputs/Travel-times/Processing/Orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = Dest_all_schools_elem_2021,
                          mode = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("2021-03-01 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          walk_speed = 3.6,
                          max_trip_duration = 30)

save("output_OD_walk_elem3", file = "data-products/OD_tt_walk_2021_3.Rdata")
rm("output_OD_walk_elem3")

# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(Orig_all_2021_4)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(Orig_all_2021_4,
                          outdir = "data-inputs/Travel-times/Processing/Orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = Dest_all_schools_elem_2021,
                          mode = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("2021-03-01 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          walk_speed = 3.6,
                          max_trip_duration = 30)

save("output_OD_walk_elem4", file = "data-products/OD_tt_walk_2021_4.Rdata")
rm("output_OD_walk_elem4")
```

### 2024-11-21 paper/Manuscript.qmd
Added 2021 schools and census to the database. Started Manuscript changes.

@@

```{r, include=FALSE}
load(file = paste0(here::here(),"/data-prep/data-products/cen_data_2011_hamiltononly.Rdata"))
cen_data_2011_hamiltononly <- cen_data_2011_1 |> filter(Region.Name == "Hamilton") #NOTE: GEOUID 35250242, 35250133, 35250457 do not have any child population
rm("cen_data_2011_1")
load(file=paste0(here::here(),"/data-prep/data-products/cen_data_2016_hamiltononly.Rdata"))
cen_data_2016_hamiltononly <- cen_data_2016_1 |> filter(Region.Name == "Hamilton") #NOTE: GEOUID 35250133, 35250398, 35250427 do not have any child population
rm("cen_data_2016_1")

ham_city_bound <- st_read(paste0(here::here(),"/data-prep/data-inputs/Boundaries/City_Boundary.shp"),
                          quiet = TRUE) |> st_transform(st_crs(cen_data_2016_hamiltononly))
ham_community_bounds <- st_read(paste0(here::here(),"/data-prep/data-inputs/Boundaries/Community_Boundaries.shp"),quiet = TRUE) |> 
  st_transform(st_crs(cen_data_2016_hamiltononly)) |> 
  mutate(COMMUNITY = case_when(COMMUNITY_ == "Ancaster" ~ "AN",
                               COMMUNITY_ == "Dundas" ~ "DU",
                               COMMUNITY_ == "Flamborough" ~ "FL",
                               COMMUNITY_ == "Glanbrook" ~ "GL",
                               COMMUNITY_ == "Hamilton" ~ "HA",
                               COMMUNITY_ == "Stoney Creek" ~ "SC"),
         COMMUNITY_ = ifelse(COMMUNITY_ == "Hamilton", "Hamilton Central", COMMUNITY_)) |> st_transform(st_crs(cen_data_2016_hamiltononly))
```

```{r, include=FALSE}
cen_data_2011_hamiltononly <- 
  cen_data_2011_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
    -c("Shape.Area", "Quality.Flags", "Type", "CD_UID", "NHS.Non.Return.Rate", "CSD_UID", "CT_UID", "CT_UID","NHS.Non.Return.Rate.1", "v_CA11N_2564..Median.after.tax.household.income..", "v_CA11F_17..15.to.19.years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA11N_2458..Median.after.tax.family.income..",
    "DA_med_AT_lone-family_income" = "v_CA11N_2476..Median.after.tax.family.income..",
    "DA_prev_LIMAT_under18" = "v_CA11N_2609..Less.than.18.years..",
    "DA_pop_5to9" = "v_CA11F_11..5.to.9.years",
    "DA_pop_10to14" = "v_CA11F_14..10.to.14.years") %>%
  mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
  dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))

cen_data_2016_hamiltononly <- 
  cen_data_2016_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
    -c("Shape.Area",  "Type", "CD_UID", "CSD_UID", "CT_UID", "CT_UID", "v_CA16_2398..Median.after.tax.income.of.households.in.2015....", "v_CA16_64..15.to.19.years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA16_2448..Median.after.tax.income.of.economic.families.in.2015....",
    "DA_med_AT_lone-family_income" = "v_CA16_2460..Median.after.tax.income.of.lone.parent.economic.families.in.2015....",
    "DA_prev_LIMAT_under18" = "v_CA16_2543..0.to.17.years....",
    "DA_pop_5to9" = "v_CA16_25..5.to.9.years",
    "DA_pop_10to14" = "v_CA16_43..10.to.14.years") %>% mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
    dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))
```

```{r, include=FALSE}
library(hamONdests) # A R package of Hamilton, Ontario destinations

data(Schools_2011_2016)
ALL_SCHOOLS_Elem <- Schools_2011_2016 %>% filter(Level == "Elementary")
rm("Schools_2011_2016")

# add community label to schools
ALL_SCHOOLS_Elem <- st_intersection(ham_community_bounds,
                        ALL_SCHOOLS_Elem |> st_transform(st_crs(cen_data_2016_hamiltononly)))

data(School_Catchments_2011_2016)
ALL_SCHOOLS_CATCH_Elem <- School_Catchments_2011_2016 %>% filter(Level == "Elementary")
rm("School_Catchments_2011_2016")

@@

```{r}
# # maybe consider using year 2021 as a comparision? 2011 vs 2016 vs 2021?
# SCHOOLS_2023 <- st_read("data-prep/data-inputs/Educational_Institutions-2023/Educational_Institutions.shp") |> filter(CATEGORY == "Elementary School" & (SCHOOL_BOA == "Hamilton-Wentworth Catholic District School Board" | SCHOOL_BOA == "Hamilton-Wentworth District School Board" )) |> st_transform(st_crs(cen_data_2011_hamiltononly))
```

```{r augmenting-catchment, include=FALSE}
ALL_SCHOOLS_CATCH_Elem_extra2011 <- ALL_SCHOOLS_CATCH_Elem |>
  left_join(ALL_SCHOOLS_Elem |> st_drop_geometry() |> group_by(CID2011) |> 
              summarize(OTGC = sum(OTGC2011, na.rm = T)), 
            by = c("CID" = "CID2011")) |> filter(!is.na(OTGC))

ALL_SCHOOLS_CATCH_Elem_extra2016 <- ALL_SCHOOLS_CATCH_Elem |>
  left_join(ALL_SCHOOLS_Elem |> st_drop_geometry() |> group_by(CID2016) |> 
              summarize(OTGC = sum(OTGC2016, na.rm = T)), 
            by = c("CID" = "CID2016")) |> filter(!is.na(OTGC))

ALL_SCHOOLS_CATCH_Elem_extra <- rbind(ALL_SCHOOLS_CATCH_Elem_extra2011, ALL_SCHOOLS_CATCH_Elem_extra2016)
rm(ALL_SCHOOLS_CATCH_Elem_extra2011,ALL_SCHOOLS_CATCH_Elem_extra2016)

#checks
sum(ALL_SCHOOLS_Elem$OTGC2011, na.rm=T) + sum(ALL_SCHOOLS_Elem$OTGC2016, na.rm=T)
ALL_SCHOOLS_CATCH_Elem_extra$OTGC |> sum(na.rm=T)

#reformating schools for mapping
ALL_SCHOOLS_Elem_nochangeorexpanded <- ALL_SCHOOLS_Elem |> filter(Year == "2011 and 2016") |> mutate(Year = ifelse(Year == "2011 and 2016", "2011", Year))
ALL_SCHOOLS_Elem_formapping <- rbind(ALL_SCHOOLS_Elem, ALL_SCHOOLS_Elem_nochangeorexpanded)
rm(ALL_SCHOOLS_Elem_nochangeorexpanded)
ALL_SCHOOLS_Elem_formapping <- ALL_SCHOOLS_Elem_formapping |> mutate(Year = ifelse(Year == "2011 and 2016", "2016", Year))
```

>>>>

<!--Chat with Antonio about the assumption regarding the 'student population' per parcel. As it stands, when the student pop for the OD matrix is summed, it equals the total student population in the city. Each parcel is assigned a DA-pop/(the number of potential parcel trips in a DA * the potential parcels trips from that parcel). This is highly correlated with we just weighted it by DA-pop/DA-dwellings (but need to double check).  --->

@@

```{r, include=FALSE}
load(paste0(here::here(),file="/data-prep/data-products/cen_data_2011_hamiltononly.Rdata"))
cen_data_2011_hamiltononly <- cen_data_2011_1 |> filter(`Region Name` == "Hamilton") #NOTE: GEOUID 35250242, 35250133, 35250457 do not have any child population
rm("cen_data_2011_1")
load(paste0(here::here(),file="/data-prep/data-products/cen_data_2016_hamiltononly.Rdata"))
cen_data_2016_hamiltononly <- cen_data_2016_1 |> filter(`Region Name` == "Hamilton") #NOTE: GEOUID 35250133, 35250398, 35250427 do not have any child population
rm("cen_data_2016_1")
load(paste0(here::here(),file="/data-prep/data-products/cen_data_2021_hamiltononly.Rdata"))
cen_data_2021_hamiltononly <- cen_data_2021_1 
rm("cen_data_2021_1")

ham_city_bound <- st_read(paste0(here::here(),"/data-prep/data-inputs/Boundaries/City_Boundary.shp"),
                          quiet = TRUE) |> st_transform(st_crs(cen_data_2016_hamiltononly))
ham_community_bounds <- st_read(paste0(here::here(),"/data-prep/data-inputs/Boundaries/Community_Boundaries.shp"),quiet = TRUE) |> 
  st_transform(st_crs(cen_data_2016_hamiltononly)) |> 
  mutate(COMMUNITY = case_when(COMMUNITY_ == "Ancaster" ~ "AN",
                               COMMUNITY_ == "Dundas" ~ "DU",
                               COMMUNITY_ == "Flamborough" ~ "FL",
                               COMMUNITY_ == "Glanbrook" ~ "GL",
                               COMMUNITY_ == "Hamilton" ~ "HA",
                               COMMUNITY_ == "Stoney Creek" ~ "SC"),
         COMMUNITY_ = ifelse(COMMUNITY_ == "Hamilton", "Hamilton Central", COMMUNITY_)) |> st_transform(st_crs(cen_data_2016_hamiltononly))

ggh_taz <- st_read(paste0(here::here(),"/data-prep/data-inputs/TTS/tts06_83_region.shp")) |> st_transform(st_crs(cen_data_2011_hamiltononly))
```

```{r, include=FALSE}
cen_data_2011_hamiltononly <- 
  cen_data_2011_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape Area", "Quality Flags", "Type", "CD_UID", "NHS Non-Return Rate", "CSD_UID", "CT_UID", "CT_UID","NHS Non Return Rate", "v_CA11N_2564: Median after-tax household income $", "v_CA11F_17: 15 to 19 years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA11N_2458: Median after-tax family income $",
    "DA_med_AT_lone-family_income" = "v_CA11N_2476: Median after-tax family income $",
    "DA_prev_LIMAT_under18" = "v_CA11N_2609: Less than 18 years %",
    "DA_pop_5to9" = "v_CA11F_11: 5 to 9 years",
    "DA_pop_10to14" = "v_CA11F_14: 10 to 14 years") %>%
  mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
  dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))

cen_data_2016_hamiltononly <- 
  cen_data_2016_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape Area",  "Type", "CD_UID", "CSD_UID", "CT_UID", "CT_UID", "v_CA16_2398: Median after-tax income of households in 2015 ($)",  "v_CA16_64: 15 to 19 years"  )) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA16_2448: Median after-tax income of economic families in 2015 ($)",
    "DA_med_AT_lone-family_income" = "v_CA16_2460: Median after-tax income of lone-parent economic families in 2015 ($)",
         "DA_prev_LIMAT_under18" = "v_CA16_2543: 0 to 17 years (%)",
         "DA_pop_5to9" = "v_CA16_25: 5 to 9 years",
         "DA_pop_10to14" = "v_CA16_43: 10 to 14 years") %>% mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
         dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))

cen_data_2021_hamiltononly <- 
  cen_data_2021_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape Area",  "Type", "CD_UID", "CSD_UID", "CT_UID", "CT_UID", "v_CA21_566: Median after-tax income in 2020 among recipients ($)" ,"v_CA21_71: 15 to 19 years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA21_966: Median after-tax income of economic family in 2020 ($)",
    "DA_med_AT_lone-family_income" = "v_CA21_978: Median after-tax income of one-parent economic families in 2020 ($)",
    "DA_prev_LIMAT_under18" =  "v_CA21_1043: 0 to 17 years",
    "DA_pop_5to9" = "v_CA21_32: 5 to 9 years",
    "DA_pop_10to14" = "v_CA21_50: 10 to 14 years") %>%
    mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
         dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))
```

```{r, include=FALSE}
# library(hamONdests) # A R package of Hamilton, Ontario destinations
load(file=paste0(here::here(),"/data-prep/data-products/SCHOOLS.Rdata")) #for this paper, the schools under analysis

# add community label to schools
SCHOOLS <- st_intersection(ham_community_bounds,
                        SCHOOLS |> st_transform(st_crs(cen_data_2016_hamiltononly)))

# data(School_Catchments_2011_2016)
# ALL_SCHOOLS_CATCH_Elem <- School_Catchments_2011_2016 %>% filter(Level == "Elementary")
# rm("School_Catchments_2011_2016")

@@

```{r augmenting-catchment, include=FALSE}
# ALL_SCHOOLS_CATCH_Elem_extra2011 <- ALL_SCHOOLS_CATCH_Elem |>
#   left_join(ALL_SCHOOLS_Elem |> st_drop_geometry() |> group_by(CID2011) |> 
#               summarize(OTGC = sum(OTGC2011, na.rm = T)), 
#             by = c("CID" = "CID2011")) |> filter(!is.na(OTGC))
# 
# ALL_SCHOOLS_CATCH_Elem_extra2016 <- ALL_SCHOOLS_CATCH_Elem |>
#   left_join(ALL_SCHOOLS_Elem |> st_drop_geometry() |> group_by(CID2016) |> 
#               summarize(OTGC = sum(OTGC2016, na.rm = T)), 
#             by = c("CID" = "CID2016")) |> filter(!is.na(OTGC))
# 
# ALL_SCHOOLS_CATCH_Elem_extra <- rbind(ALL_SCHOOLS_CATCH_Elem_extra2011, ALL_SCHOOLS_CATCH_Elem_extra2016)
# rm(ALL_SCHOOLS_CATCH_Elem_extra2011,ALL_SCHOOLS_CATCH_Elem_extra2016)
# 
# #checks
# sum(ALL_SCHOOLS_Elem$OTGC2011, na.rm=T) + sum(ALL_SCHOOLS_Elem$OTGC2016, na.rm=T)
# ALL_SCHOOLS_CATCH_Elem_extra$OTGC |> sum(na.rm=T)

#reformatting schools for mapping
SCHOOLS_nochangeorexpanded <- SCHOOLS |> filter(Year == "2011 and 2016" & Year_2 == "2016 and 2021") |> 
  mutate(Year = ifelse(Year == "2011 and 2016", "2011", Year))
ALL_SCHOOLS_Elem_formapping <- rbind(SCHOOLS, SCHOOLS_nochangeorexpanded)
rm(SCHOOLS_nochangeorexpanded)
ALL_SCHOOLS_Elem_formapping <- ALL_SCHOOLS_Elem_formapping |> mutate(Year = ifelse(Year == "2011 and 2016", "2016", Year))
```
